\documentclass[fleqn,11pt]{article}
\usepackage[includeheadfoot,margin=0.5in,headheight=18pt]{geometry}
\usepackage[english]{babel}
% \usepackage{firamath-otf}
% \usepackage{cmbright}
% \usepackage[OT1]{fontenc}
\usepackage[sfdefault]{notomath}
\usepackage{calc}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{multicol}
\usepackage[table,xcdraw]{xcolor}
\usepackage[version=4]{mhchem}
\usepackage[sticky-per,print-unity-mantissa=false]{siunitx}
\usepackage{cancel}
\usepackage{contour}
\usepackage{ulem}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{arydshln}
\usepackage{pdfpages}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{wrapfig}
\usepackage{makecell}
\usepackage{float}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{circuitikz}
\usepackage{tikzpagenodes}
% \usepackage{tabularx}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{parskip}
\usepackage{setspace}
\usepackage{cleveref}
% \setoperatorfont\mathsf
% \renewcommand*\familydefault{\sfdefault}

\frenchspacing
\renewcommand{\ULdepth}{1.8pt}
\contourlength{0.8pt}

\usetikzlibrary{positioning}
\usepgfplotslibrary{external}
% \tikzexternalize{}

\pgfplotsset{compat=1.18}
\def\axisdefaultwidth{5in}

\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  breakatwhitespace=false,
  numbers=left,
  keywordstyle=\pmb,
  numberstyle=\tiny,
  commentstyle=\itshape,
  showstringspaces=false,
}

\renewcommand{\title}{Homework 9} % TODO change this

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  filecolor=magenta,      
  urlcolor=cyan,
  pdftitle={\title},
  pdfpagemode=FullScreen,
}

\pagestyle{fancyplain}
\lhead{\Large \textbf{\title}}
\rhead{\large Sumit Basak}
\lfoot{NERS 561}
\cfoot{}
\rfoot{Page \thepage~of~\pageref{LastPage}}

\newcommand{\md}[2][d]{\ensuremath{\operatorname{#1}\!{#2}}}
\newcommand{\mder}[3][d]{\frac{\md[#1]{#2}}{\md[#1]{#3}}}
\newcommand{\textarrow}[2]{\xRightarrow{\substack{\text{#1} \\ \text{#2}}}}
\renewcommand{\vec}[1]{\ensuremath{\mathbf{#1}}}

\sisetup{exponent-product=*}
\colorlet{darkgreen}{green!50!black}

\DeclareSIUnit{\year}{yr}
\DeclareSIUnit{\curie}{Ci}
\DeclareSIUnit{\calorie}{cal}
\DeclareSIUnit{\atom}{atom}
\DeclareSIUnit{\df}{\SIUnitSymbolDegree F}
\DeclareSIUnit{\btu}{BTU}
\DeclareSIUnit{\lbm}{lbm}
\DeclareSIUnit{\inch}{in}
\DeclareSIUnit{\atom}{atoms}
\DeclareSIUnit{\barn}{b}

\newcommand{\textover}[3][c]{%
  % #1 is the alignment, default c
  % #2 is the text to be printed
  % #3 is the text for setting the width
  \makebox[\widthof{#3}][#1]{#2}%
}

\newcommand{\myuline}[1]{%
  \uline{\phantom{#1}}%
  \llap{\contour{white}{#1}}%
}

\begin{document}
\section{Implementation}
The code for this assignment is attached in appendix. It is 
split into seven regions, marked with \texttt{\#region 
\(\dots\) \#endregion} statements. A description of each 
region follows. \begin{description}
  \item [Input] Parses the given input file for the required parameters.

  \item [Mesh Balance Equations] Sets up four vectors, \(a\) through \(d\), with the following contents, where \(i\) is the mesh index.
    \begin{align*}
      a_i &= -\tilde D_{i - 1} \\
      b_i &= \Sigma_{a, i} \md{h} - a_i - c_i \\
      c_i &= -\tilde D_i \\
      d_i &= \nu \Sigma_{f, i} \md{h}
      \intertext{where,}
      \tilde D_i &= \frac{1}{\frac{\md{h}}{2 D_i} + \frac{\md{h}}{2 D_{i + 1}}}
    \end{align*}
  
  \item [LU Factorization] Splits the transformation matrix \(M\) into \(L\) and \(U\) by creating vectors \(\tilde a\) and \(\tilde b\) as follows.
    Also prepares \(\psi = \lambda F \phi\), the fission source in each mesh region.
    \begin{align*}
      \tilde a &= \frac{a_i}{\tilde b_{i - 1}} \\
      \tilde b &= b_i - \tilde a_i c_{i - 1}
      \intertext{The matricies are defined as,}
      L &= \begin{bmatrix}
        1 & 0 & 0 & 0 & 0 \\
        \tilde a_2 & 1 & \ddots & \ddots & 0 \\
        0 & \ddots & \ddots & \ddots & 0 \\
        0 & \ddots & \ddots & \ddots & 0 \\
        0 & 0 & 0 & \tilde a_N & 1
      \end{bmatrix} \\
      U &= \begin{bmatrix}
        \tilde b_1 & c_1 & 0 & 0 & 0 \\
        0 & \tilde b_2 & \ddots & \ddots & 0 \\
        0 & \ddots & \ddots & \ddots & 0 \\
        0 & \ddots & \ddots & \ddots & c_{N - 1} \\
        0 & 0 & 0 & 0 & \tilde b_N
      \end{bmatrix}
    \end{align*}
  
  \item [Forward Elimination] Obtains a temporary vector \(y\) by solving \(L y = \psi\) through forward elimination.
  
  \item [Backward Substitution] Obtains the flux by solving \(U \phi = y\) by backward substitution.
  
  \item [Outer Iteration] Checks for convergence in the flux \(\phi\) and eigenvalue \(\lambda\) and loops if needed.
  
  \item [Output] Normalizes flux and fission source and prints them to a file.
\end{description}

\section{Results}
We are told to solve find the flux and fission source given 
the following input file. This input specifies a \SI{180}{\centi\meter}
problem domain with nine regions and five meshes per region.
There are four possible composition each region can take,
and their mappings are provided on Line~10. Tolerance for 
\(k\) is \num{e-6}, while that of flux is \num{e-5}. We use 
the infinity norm -- that is, the maximum absolute difference 
among the elements of the vector -- for the latter. Results 
are plotted below.

\lstinputlisting{../input-180.txt}

The resulting eigenvalue is \(k = 0.87309\), with a dominance
ratio \(\sigma = 0.43611\). This solution took 175 iterations
to converge. In general, we see an expected shape of the 
flux distribution. Local maxima and minima correspond to 
regions of increased and decreased diffusion and fission 
cross section, respectively. Region Six, from 
\SIrange{100}{120}{\centi\meter}, has as lower average 
magnitude than that of Region Four, from 
\SIrange{60}{80}{\centi\meter}, matching the input file. 
This is most likely due to the decreased fission cross 
section.

\begin{center}
  \begin{tikzpicture}[
    trim axis left,
    trim axis right,
  ]
    \begin{axis}[
      xlabel={\(x\) {[\si{\centi\meter}]}},
      ylabel={Normalized Value},
      legend pos={outer north east},
      legend cell align={left},
      ytick distance={1e-2}
    ]
      \addplot[mark=none,const plot,solid] table {../phi-180-noshift.txt};
      \addplot[mark=none,const plot,dotted] table {../source-180-noshift.txt};

      \addlegendentry{Flux}
      \addlegendentry{Fission Source}
    \end{axis}
  \end{tikzpicture}
\end{center}

\newpage 
\section{Wielandt Shift}
\subsection{Doubled Domain}
We first rerun the program unchanged, doubling the domain 
size to \SI{360}{\centi\meter}. In this case, the eigenvalue
is \(k = 0.93583\) and the dominance ratio is \(\sigma = 
0.64411\). This time, the solution took 1790 iterations to 
converge, an order of magnitude increase. The effects of 
the lower fission cross section in Region Six (this time 
from \SIrange{200}{240}{\centi\meter}) are magnified
in this larger domain.

\begin{center}
  \begin{tikzpicture}[
    trim axis left,
    trim axis right,
  ]
    \begin{axis}[
      xlabel={\(x\) {[\si{\centi\meter}]}},
      ylabel={Normalized Value},
      legend pos={outer north east},
      legend cell align={left},
      ymax={0.09999},
      xtick distance={40},
    ]
      \addplot[mark=none,const plot,solid] table {../phi-360-noshift.txt};
      \addplot[mark=none,const plot,dotted] table {../source-360-noshift.txt};

      \addlegendentry{Flux}
      \addlegendentry{Fission Source}
    \end{axis}
  \end{tikzpicture}
\end{center}

\subsection{Shift Implementation}
By offsetting the eigenvalue by a certain amount, we can 
achieve faster convergence. This offset is defined as 
\(\lambda_s\), where \(\tilde \lambda = \lambda - \lambda_s\).
We apply this shift after some arbitrary number of iterations.
The table below summarizes convergence times for a few of 
these skip values.
Note that we are not modifying the problem itself, just the 
convergence time. So, the solution (flux, fission source, 
and eigenvalue) is all identical for all the shift values.

\begin{center}
  \begin{tabular}{c S[table-format=1.5] c}
    \toprule
    \textbf{Skipped Iterations} & \textbf{Dominance Ratio} & \textbf{Convergence Time} \\ \midrule
    2 & 0.64411 & 278 \\
    5 & 0.67474 & 278 \\
    10 & 0.85231 & 282 \\
    \(\infty\) (no shift) & 0.93583 & 1790 \\ \bottomrule
  \end{tabular}
\end{center}

We see that applying the Wielandt shift does indeed help 
significantly with convergence time. There seems to be 
diminishing returns with respect to when to start the 
shifting, but the execution time is not affected, so it does 
not really matter.

\newpage \appendix 
\section{Source Code}
The full source code for both the solver and this report is 
available at \url{https://github.com/mittsq/ners-561-hw9}; 
I made sure to make the repository public this time. The 
important parts of the code are included below.

\lstinputlisting[language={[Sharp]C}]{../code/Program.cs}
\end{document}
